Precision calculations of the cosmic shear power spectrum projection

### 2017-02-16 limber2.tex
MK: Replaced Legendre transform with spherical transform; added ref Peebles 73

For future
large-scale surveys with unprecedented precision, we show that the spherical
second-order Limber approximation will provide sufficient accuracy. \ch{In this case, the
cosmic-shear power spectrum is shown to be in agreement with the full projection at the
sub-percent level for $\ell > 3$, with the corresponding errors an order of
magnitude below cosmic variance for all $\ell$.
%We show that the commonly used 
%shear two-point correlation function is accurate well below one percent up to a few hundred arc
%minutes when adopting. 
%We find that when using the fast Hankel transformation, formally defined only in flat
%%coordinates, to compute the commonly-used two-point function from the power spectrum results
%in errors below two percent compared to the spherical Legendre transformation.}
When computing the two-point shear correlation function, we show that the flat-sky fast Hankel 
transformation results in errors below two percent compared to the full spherical Legendre transformation.}
%
In the spirit of reproducible research, our numerical implementation of all
approximations and the full projection are publicly available within the
package \textsc{nicaea} at \texttt{http://www.cosmostat.org/software/nicaea}.

@@

We further show that the second-order Limber approximation is an
accurate representation of the full projection, with better than percent level
precision for scales $\ell > 3$. Since this approximation involves only 1D
integrals over the matter power spectrum, it is very fast to calculate
numerically and can readily be employed in Monte-Carlo sampling methods to
obtain precision constraints on cosmological parameters.
\mk{We also compute the shear correlation function using a spherical Legendre
transformation, and compare this to the flat-sky approximated, commonly used
fast Hankel transformation.}

@@

The lensing potential $\psi$ at a position on the sky $(\theta, \varphi)$ in
the Born approximation is defined as the projected 3D metric potential $\Phi$
along the line of sight of a flat Universe \citep{1998ApJ...498...26K,BS01},
%
% Hu00 (21,22) with minus sign
%
\begin{equation}
  \psi(\theta, \varphi) = \frac 2 {{\rm c}^2} \int\limits_0^\infty \frac{{\rm d}\chi}{\chi}
    %\Phi[\chi, f_K(\chi) \theta, f_K(\chi) \varphi; \chi] \, q(\chi),
    \Phi[\chi, \chi \theta, \chi \varphi; \chi] \, q(\chi),
  \label{eq:psi}
\end{equation}
%
where the lensing efficiency $q$ is given by
%
\begin{equation}
  q(\chi) = \int\limits_\chi^{\chi_{\rm h}} {\rm d} \chi^\prime \, n(\chi^\prime)
    %\frac{f_K(\chi^\prime - \chi)}{f_K(\chi^\prime)},
    \frac{\chi^\prime - \chi}{\chi^\prime},
  %
  \label{eq:lens_efficiency}
\end{equation}
%
corresponding to a population of lensed galaxies with a normalised source
redshift distribution $n_z(z) {\rm d}z = n(\chi) {\rm d} \chi$, with the limit
being the comoving distance to the horizon $\chi_{\rm h}$\footnote{\mk{In
(\ref{eq:psi}) we have replaced this limit without loss of generality with
$\infty$, since $q(\chi > \chi_{\rm h}) = 0$.}}. Here, ${\rm c}$ is the speed of
light, and the projection is carried out over comoving distances $\chi$.
%, and $f_K(\chi) = |K|^{-1/2} S_K(|K|^{1/2} \chi)$ is the comoving angular
% diameter distance, where the function $S_K$ equals $\sin$, ${\rm id}$, or
% $\sinh$ if the spatial curvature of the Universe $K$ is positive, zero, or
% negative, respectively.
The last argument of the potential $\Phi$ is not to
be understood as coordinate, but rather as a substitute of cosmic time,
$t(\chi)$, to express the time at which the potential is evaluated. This is
true in the following for all fields and functions thereof that dynamically
change with cosmic epoch.

@@

Following \cite{2000PhRvD..62d3007H} we decompose the lensing
potential (equation~\ref{eq:psi}) into spherical harmonics, in analogy
to the cosmic microwave background (CMB)
temperature, both of which are scalar functions on the sphere. This
decomposition and its inverse are
%
% Hu00 (23)
%
\begin{align}
  \psi(\theta, \varphi) = \sum_{\ell=0}^\infty \sum_{m=-\ell}^\ell \psi_{\ell m} {\rm Y}_{\ell m}(\theta, \varphi);
    \label{eq:psi_harm_exp}
    \\
    %
  \psi_{\ell m} = \int_{\mathbb{S}^2} {\rm d} \Omega \, \psi(\theta, \varphi) {\rm Y}^\ast_{\ell m}(\theta, \varphi).
  \label{eq:psi_harm_exp_inv}
\end{align}
%
Complex conjugation is denoted with the superscript $^\ast$.
To specify tomographic redshift bins $i=1\ldots N_z$, we introduce a family
of lensing efficiency functions $q_i$ defined by a corresponding family of
redshift distributions $n_i$ via equation~(\ref{eq:lens_efficiency}). The resulting
lensing potential is denoted by $\psi_{\ell m, i}$. The tomographic power
spectrum of the lensing potential between two redshift bins $i$ and $j$,
$C_{ij}^\psi$ \citep{pee80} is then defined as
%
\begin{equation}
  \left\langle \psi^{}_{\ell m, i} \, \psi^\ast_{\ell^\prime m^\prime, j} \right\rangle
    = \delta_{\ell \ell^\prime} \delta_{m m^\prime} C^\psi_{ij}(\ell) .
  \label{eq:C_ell_psi}
\end{equation}
%
\mk{Note that the argument $\ell$ is a discrete integer variable, and is often written as index, $C_\ell$.}
Using the properties of the spherical harmonics (see
App.~\ref{sec:derivations_C} for details) the power spectrum can be written as
%
\begin{align}
  C_{ij}^\psi(\ell) = & \frac 8 {c^4 \pi} 
  \int_0^\infty \frac{{\rm d}\chi}{\chi} q_i(\chi)
  \int_0^\infty \frac{{\rm d}\chi^\prime}{\chi^\prime} q_j(\chi^\prime)
  \int {\rm d} k k^2 \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_\Phi(k; \chi, \chi^\prime)
  \label{eq:C_ell_phi_Pphi} \\
  = & \frac 8 \pi \pref^2
  \int_0^\infty \frac{{\rm d}\chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
  \int_0^\infty \frac{{\rm d}\chi^\prime}{\chi^\prime} \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
  \int \frac{{\rm d} k}{k^2} \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_{\rm m}(k; \chi, \chi^\prime) ;
  \label{eq:C_ell_phi_Pm}
\end{align}
%
where ${\rm j}_\ell$ is the spherical Bessel function of order $\ell$. For
convenience we introduce the normalisation constant $\pref$ as
%
\begin{equation}
  \pref = \frac 3 2 \Omega_{\rm m} \left(\frac{H_0} c\right)^2.
  \label{eq:pref}
\end{equation}
%
The first line expresses $C_{ij}^\psi$ in terms of the 3D potential power spectrum $P_\Phi$, defined as
%
\begin{align}
  \left\langle \hat \Phi(\vec k; \chi) \hat \Phi^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_\Phi(k; \chi, \chi^\prime).
    %\nonumber \\
    %= & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) \pref^2 k^{-4} a^{-1}(\chi) a^{-1}(\chi^\prime)
      %P_{\rm m}(k; \chi, \chi^\prime).
  \label{eq:p_phi}
\end{align}
%
The second line uses the 3D matter density power spectrum $P_{\rm m}$, which is defined analogously as
%
\begin{align}
  \left\langle \hat \delta(\vec k; \chi) \hat \delta^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_{\rm m}(k; \chi, \chi^\prime),
  \label{eq:p_m}
\end{align}
%
and is related to $P_\Phi$ via the absolute square of the Poisson equation
(\ref{eq:poisson}).

\mk{This type of cross-power spectrum between different cosmological epochs
$\chi$ and $\chi^\prime$ was introduced in \cite{2005PhRvD..72b3516C}. In
Sects.~\ref{sec:geom_mean} and \ref{sec:factor_D} this unequal-time
cross-spectrum \citep{2016arXiv161200770K} will be further evaluated and
simplified in the context of the Limber approximation. The oscillating Bessel
functions in equation~(\ref{eq:C_ell_phi_Pm}) ensure that only relatively close epochs
contribute to the lensing potential correlation. This make sense since observed
light rays from two galaxies at different positions on the sky that necessarily
converge at the observer today, pick up the density fluctuations at similar
times while propagating through the large-scale structure. A similar argument
has been made in \cite{BS01}: since the matter power spectrum scales with $k$
for $k \rightarrow 0$, there is decreasing power towards larger and larger
scales. In particular, the correlation of cosmic fields decreases strongly
above a coherence scale $|\chi - \chi^\prime| \gsim L_{\rm coh}$ which is significantly
smaller than the horizon scale $\chi_{\rm h}$. }

@@

The majority of cosmic shear analyses have used the predicted lensing power
spectrum approximated on a flat sky, neglecting the sky curvature. This is a
valid approach for past and current survey areas with an extent less than 10
degrees. To account for the sky curvature of the observed data, the shear
correlation functions from observed galaxy ellipticities is now routinely
computed using spherical coordinates, since projecting to a Cartesian plane has
been shown to cause significant biases of the two-point funtion on large scales
\citep{FSHK08}, \mk{and lead to spurious B-modes \citep{asgari/etal:2017}}. Here
we examine the effect of sky curvature on the theoretical models of the power
spectrum, and the effect on cosmological parameter inference (see
Sect.~\ref{sec:cfhtlens}).

@@

For a spin-2 field, \cite{2000PhRvD..62d3007H} approximates the edth operator
by Cartesian derivatives, and approximates equation~(\ref{eq:sYlm_def}) as
%
\begin{equation}
  \ell^2 \, _{\pm2}\!{\rm Y}_{\ell m}(\theta, \varphi) \approx  {\rm e}^{{\mp}2 {\rm i} \phi_\ell}
    (\partial_1 \pm {\rm i} \partial_2)^2 \, {\rm Y}_{\ell m}(\theta, \varphi).
  \label{eq:sYlm_def_flat}
\end{equation}
%
The spin pre-factor $\ellbar(\ell, 2) = \sqrt{(\ell-1) \ell (\ell+1) (\ell+2)}$
is replaced by $\ell^2$ in flat co-ordinates, an approximation that holds for
large $\ell$, \mk{since sky curvature can be neglected for small angular
scales.} We find for the flat-sky shear power spectrum
%
\begin{align}
  P^\gamma_{ij}(\ell) 
                 = & \frac 2 \pi \, \ell^4 \, \pref^2
                 \int_{0}^\infty \frac{\rm d \chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
                \int_{0}^\infty \frac{\rm d \chi^\prime}{\chi^\prime}
                \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
                %\nonumber \\
                %& \times
                \int_0^\infty \frac{{\rm d} k}{k^2} \, P_{\rm m}(k, \chi, \chi^\prime) \,
                {\rm j}_\ell(k \chi) \, {\rm j}_\ell(k \chi^\prime) .
  \label{eq:P_ell_gamma_full_der}
\end{align}
%
with flat-sky pre-factor $\ell^4$. See App.~\ref{sec:schmidt08} and
\ref{sec:giannantonio12} for discussions of alternative expressions for the
flat-sky power spectrum.

@@

To separate the $k$- and $\chi, \chi^\prime$-terms in
equation~(\ref{eq:C_ell_full}), we first approximate the matter power
cross-spectrum between two distances by the geometric mean of the two involved
distances \citep{2005PhRvD..72b3516C,2016arXiv161200770K},
%
\begin{equation}
 P_{\rm m}(k; \chi, \chi^\prime) = \sqrt{ P_{\rm m}(k; \chi) P_{\rm m}(k; \chi^\prime) } \, .
  \label{eq:P_geom_mean}
\end{equation}
%
\mk{This form is justified when considering the linear power spectrum, and
follows when inserting the linearly evolving density contrast $\hat \delta(\vec
k; \chi) = D_+(\chi) \hat \delta_0(\vec k)$ into equation~(\ref{eq:p_m}), where
$\delta_0$ is the present-day linearly extrapolated density contrast, and $D_+$
the linear growth factor with $D_+(0) = 1$. This is a good approximation
also in the non-linear case as shown in \cite{2016arXiv161200770K}.}

@@

Following \cite{2008PhRvD..78l3506L} we expand to third order
%
\begin{equation}
  \lim_{\varepsilon \rightarrow 0} \int_0^\infty {\rm d} x \, {\rm e}^{-\epsilon (x - \nu)} g(x) {\rm J}_\nu(x)
  %\approx g(\nu) - \frac 1 2 g^{\prime\prime}(\nu) - \frac \nu 6 g^{\prime\prime\prime}(\nu) 
  \approx g(\nu) - \frac 1 2 \left.\frac{{\rm d}^2 g}{{\rm d} x^2}\right|_{x=\nu}
                 - \frac \nu 6 \left.\frac{{\rm d}^3 g}{{\rm d} x^3}\right|_{x=\nu},
\end{equation}
%
with $g(x) = k^{-1} f(k, \chi)$, $\chi=x/k$, and its derivatives $g^{(n)}(x) =
k^{-1-n} f^{(n)}(k, \chi)$ for a given $k$, where the derivatives of $f$
are with respect to the second argument $\chi$. \mk{In this series expansion, the
replacement of the integral with the evaluation of $g$ and its derivatives at
the maximum of the Bessel function is a good approximation of the integral if
$g$ is varying more slowly than the oscillating Bessel function.
% [Check this].
As we will show below, $f$ is a slowly varying function of the comoving
distance.} In our case the projection kernel is
%
\begin{equation}
  f(k, \chi) = \sqrt{P_{\rm m}(k; \chi)} \, a^{-1}(\chi) \chi^{-3/2} q(\chi).
  \label{eq:f_LA08}
\end{equation}
%
In the tomographic case, the index $i$ is added to $q$ and $f$.
%
Replacing both distance integrals in equation~(\ref{eq:C_ell_full_Pk}) by their
Taylor-expansions around the maxima $\nu(\ell) = \ell + 1/2$ of the two Bessel
functions, which are $k \chi$ and $k \chi^\prime$, respectively, yields
%
\begin{align}
  C^\gamma_{ij}(\ell) \approx & \, \ellbar^2(\ell, 2) \, \pref^2
    \int_0^\infty \frac{{\rm d} k}{k^3} \, k^{-2}
    \left[ f_i(k, \chi) - \frac{1}{2 k^2} f_i^{\prime\prime}(k, \chi)
      - \frac{\nu(\ell)}{6 k^3} f_i^{\prime\prime\prime}(k, \chi) + \ldots \right]
    %\nonumber \\
    %& \times
    \left[ f_j(k, \chi) - \frac{1}{2 k^2} f_j^{\prime\prime}(k, \chi)
    - \frac{\nu(\ell)}{6 k^3} f_j^{\prime\prime\prime}(k, \chi) + \ldots \right] \, .
  \label{eq:C_ell_limber2_dk}
\end{align}
%
Changing the integration to $\chi = \nu/k$ and collecting terms according to their $\nu$-dependence:
%
\begin{align}
  C^\gamma_{ij}(\ell) \approx & \, C^\gamma_{{\rm L1}, ij}(\ell) + C^\gamma_{{\rm L2}, ij}(\ell)
      %\nonumber \\
    = %&
    \frac{\ellbar^2(\ell, 2)}{\nu^4(\ell)} \, \pref^2
    \int_0^\infty {\rm d} \chi \, \chi^3 \, % P_{\rm m}\left( k \right)
    \Bigg\{
    %\left\{
    (f_i f_j)(\nu(\ell)/\chi, \chi)
    %\right.
      \nonumber \\
    &
    %\left.
     - \frac 1 {\nu^2(\ell)}
    \left[ \frac{\chi^2}{2} \left( f^{}_i f^{\prime\prime}_j + f_i^{\prime\prime} f^{}_j \right)(\nu(\ell)/\chi, \chi)
     + \frac{\chi^3}{6} \left( f^{}_i f^{\prime\prime\prime}_j + f^{\prime\prime\prime}_i f^{}_j \right)(\nu(\ell)/\chi, \chi)
    \right]
    + {\cal O}(\nu^{-4})
    %\right\}
    \Bigg\}
    \, .
  \label{eq:C_ell_limber12_dr}
\end{align}
%
The first term corresponds to the well-known first-order Limber approximation
\citep{1953ApJ...117..134L,1992ApJ...388..272K}, which is widely used in
weak gravitational lensing. We retrieve the (spherical) standard expression by
inserting back the projection kernel (equation~\ref{eq:f_LA08}),
%
\begin{align}
  C^\gamma_{{\rm L1}, ij}(\ell) = & \, \frac{\ellbar^2(\ell, 2)}{\nu^4(\ell)} \, \pref^2 \int {\rm d} \chi \frac{ q_i(\chi) q_j(\chi) }{a^2(\chi)}
  P_{\rm m}\left(\frac{\nu(\ell)}{\chi}; \chi\right) . \Label{ExtL1Sph}
  \label{eq:C_ell_limber1}
\end{align}
%
In the Limber approximation, modes between structures at different epochs do
not contribute to the single line-of-sight integration.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{\mk{Approximation for small $\ell$}}
\label{sec:factor_D}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

For future
large-scale surveys with unprecedented precision, we show that the spherical
second-order Limber approximation will provide sufficient accuracy. \ch{In this case, the
cosmic-shear power spectrum is shown to be in agreement with the full projection at the
sub-percent level for $\ell > 3$, with the corresponding errors an order of
magnitude below cosmic variance for all $\ell$.
When computing the two-point shear correlation function, we show that the flat-sky fast Hankel 
transformation results in errors below two percent compared to the full spherical transformation.}
%
In the spirit of reproducible research, our numerical implementation of all
approximations and the full projection are publicly available within the
package \textsc{nicaea} at \texttt{http://www.cosmostat.org/software/nicaea}.

@@

We further show that the second-order Limber approximation is an
accurate representation of the full projection, with better than percent level
precision for scales $\ell > 3$. Since this approximation involves only 1D
integrals over the matter power spectrum, it is very fast to calculate
numerically and can readily be employed in Monte-Carlo sampling methods to
obtain precision constraints on cosmological parameters.
We also compute the shear correlation function using a spherical 
transformation, and compare this to the flat-sky approximated, commonly used
fast Hankel transformation.

@@

The lensing potential $\psi$ at a position on the sky $(\theta, \varphi)$ in
the Born approximation is defined as the projected 3D metric potential $\Phi$
along the line of sight of a flat Universe \citep{1998ApJ...498...26K,BS01},
%
% Hu00 (21,22) with minus sign
%
\begin{equation}
  \psi(\theta, \varphi) = \frac 2 {{\rm c}^2} \int\limits_0^\infty \frac{{\rm d}\chi}{\chi}
    %\Phi[\chi, f_K(\chi) \theta, f_K(\chi) \varphi; \chi] \, q(\chi),
    \Phi[\chi, \chi \theta, \chi \varphi; \chi] \, q(\chi),
  \label{eq:psi}
\end{equation}
%
where the lensing efficiency $q$ is given by
%
\begin{equation}
  q(\chi) = \int\limits_\chi^{\chi_{\rm h}} {\rm d} \chi^\prime \, n(\chi^\prime)
    %\frac{f_K(\chi^\prime - \chi)}{f_K(\chi^\prime)},
    \frac{\chi^\prime - \chi}{\chi^\prime},
  %
  \label{eq:lens_efficiency}
\end{equation}
%
corresponding to a population of lensed galaxies with a normalised source
redshift distribution $n_z(z) {\rm d}z = n(\chi) {\rm d} \chi$, with the limit
being the comoving distance to the horizon $\chi_{\rm h}$\footnote{In
(\ref{eq:psi}) we have replaced this limit without loss of generality with
$\infty$, since $q(\chi > \chi_{\rm h}) = 0$.}. Here, ${\rm c}$ is the speed of
light, and the projection is carried out over comoving distances $\chi$.
%, and $f_K(\chi) = |K|^{-1/2} S_K(|K|^{1/2} \chi)$ is the comoving angular
% diameter distance, where the function $S_K$ equals $\sin$, ${\rm id}$, or
% $\sinh$ if the spatial curvature of the Universe $K$ is positive, zero, or
% negative, respectively.
The last argument of the potential $\Phi$ is not to
be understood as coordinate, but rather as a substitute of cosmic time,
$t(\chi)$, to express the time at which the potential is evaluated. This is
true in the following for all fields and functions thereof that dynamically
change with cosmic epoch.

@@

Following \cite{2000PhRvD..62d3007H} we decompose the lensing
potential (equation~\ref{eq:psi}) into spherical harmonics, in analogy
to the cosmic microwave background (CMB)
temperature, both of which are scalar functions on the sphere. This
decomposition and its inverse are
%
% Hu00 (23)
%
\begin{align}
  \psi(\theta, \varphi) = \sum_{\ell=0}^\infty \sum_{m=-\ell}^\ell \psi_{\ell m} {\rm Y}_{\ell m}(\theta, \varphi);
    \label{eq:psi_harm_exp}
    \\
    %
  \psi_{\ell m} = \int_{\mathbb{S}^2} {\rm d} \Omega \, \psi(\theta, \varphi) {\rm Y}^\ast_{\ell m}(\theta, \varphi).
  \label{eq:psi_harm_exp_inv}
\end{align}
%
Complex conjugation is denoted with the superscript $^\ast$.
To specify tomographic redshift bins $i=1\ldots N_z$, we introduce a family
of lensing efficiency functions $q_i$ defined by a corresponding family of
redshift distributions $n_i$ via equation~(\ref{eq:lens_efficiency}). The resulting
lensing potential is denoted by $\psi_{\ell m, i}$. The tomographic power
spectrum of the lensing potential between two redshift bins $i$ and $j$,
$C_{ij}^\psi$ \citep{pee80} is then defined as
%
\begin{equation}
  \left\langle \psi^{}_{\ell m, i} \, \psi^\ast_{\ell^\prime m^\prime, j} \right\rangle
    = \delta_{\ell \ell^\prime} \delta_{m m^\prime} C^\psi_{ij}(\ell) .
  \label{eq:C_ell_psi}
\end{equation}
%
Note that the argument $\ell$ is a discrete integer variable, and is often written as index, $C_\ell$.
Using the properties of the spherical harmonics (see
App.~\ref{sec:derivations_C} for details) the power spectrum can be written as
%
\begin{align}
  C_{ij}^\psi(\ell) = & \frac 8 {c^4 \pi} 
  \int_0^\infty \frac{{\rm d}\chi}{\chi} q_i(\chi)
  \int_0^\infty \frac{{\rm d}\chi^\prime}{\chi^\prime} q_j(\chi^\prime)
  \int {\rm d} k k^2 \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_\Phi(k; \chi, \chi^\prime)
  \label{eq:C_ell_phi_Pphi} \\
  = & \frac 8 \pi \pref^2
  \int_0^\infty \frac{{\rm d}\chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
  \int_0^\infty \frac{{\rm d}\chi^\prime}{\chi^\prime} \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
  \int \frac{{\rm d} k}{k^2} \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_{\rm m}(k; \chi, \chi^\prime) ;
  \label{eq:C_ell_phi_Pm}
\end{align}
%
where ${\rm j}_\ell$ is the spherical Bessel function of order $\ell$. For
convenience we introduce the normalisation constant $\pref$ as
%
\begin{equation}
  \pref = \frac 3 2 \Omega_{\rm m} \left(\frac{H_0} c\right)^2.
  \label{eq:pref}
\end{equation}
%
The first line expresses $C_{ij}^\psi$ in terms of the 3D potential power spectrum $P_\Phi$, defined as
%
\begin{align}
  \left\langle \hat \Phi(\vec k; \chi) \hat \Phi^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_\Phi(k; \chi, \chi^\prime).
    %\nonumber \\
    %= & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) \pref^2 k^{-4} a^{-1}(\chi) a^{-1}(\chi^\prime)
      %P_{\rm m}(k; \chi, \chi^\prime).
  \label{eq:p_phi}
\end{align}
%
The second line uses the 3D matter density power spectrum $P_{\rm m}$, which is defined analogously as
%
\begin{align}
  \left\langle \hat \delta(\vec k; \chi) \hat \delta^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_{\rm m}(k; \chi, \chi^\prime),
  \label{eq:p_m}
\end{align}
%
and is related to $P_\Phi$ via the absolute square of the Poisson equation
(\ref{eq:poisson}).

This type of cross-power spectrum between different cosmological epochs
$\chi$ and $\chi^\prime$ was introduced in \cite{2005PhRvD..72b3516C}. In
Sects.~\ref{sec:geom_mean} and \ref{sec:factor_D} this unequal-time
cross-spectrum \citep{2016arXiv161200770K} will be further evaluated and
simplified in the context of the Limber approximation. The oscillating Bessel
functions in equation~(\ref{eq:C_ell_phi_Pm}) ensure that only relatively close epochs
contribute to the lensing potential correlation. This make sense since observed
light rays from two galaxies at different positions on the sky that necessarily
converge at the observer today, pick up the density fluctuations at similar
times while propagating through the large-scale structure. A similar argument
has been made in \cite{BS01}: since the matter power spectrum scales with $k$
for $k \rightarrow 0$, there is decreasing power towards larger and larger
scales. In particular, the correlation of cosmic fields decreases strongly
above a coherence scale $|\chi - \chi^\prime| \gsim L_{\rm coh}$ which is significantly
smaller than the horizon scale $\chi_{\rm h}$.

@@

The majority of cosmic shear analyses have used the predicted lensing power
spectrum approximated on a flat sky, neglecting the sky curvature. This is a
valid approach for past and current survey areas with an extent less than 10
degrees. To account for the sky curvature of the observed data, the shear
correlation functions from observed galaxy ellipticities is now routinely
computed using spherical coordinates, since projecting to a Cartesian plane has
been shown to cause significant biases of the two-point funtion on large scales
\citep{FSHK08}, and lead to spurious B-modes \citep{asgari/etal:2017}. Here
we examine the effect of sky curvature on the theoretical models of the power
spectrum, and the effect on cosmological parameter inference (see
Sect.~\ref{sec:cfhtlens}).

@@

For a spin-2 field, \cite{2000PhRvD..62d3007H} approximates the edth operator
by Cartesian derivatives, and approximates equation~(\ref{eq:sYlm_def}) as
%
\begin{equation}
  \ell^2 \, _{\pm2}\!{\rm Y}_{\ell m}(\theta, \varphi) \approx  {\rm e}^{{\mp}2 {\rm i} \phi_\ell}
    (\partial_1 \pm {\rm i} \partial_2)^2 \, {\rm Y}_{\ell m}(\theta, \varphi).
  \label{eq:sYlm_def_flat}
\end{equation}
%
The spin pre-factor $\ellbar(\ell, 2) = \sqrt{(\ell-1) \ell (\ell+1) (\ell+2)}$
is replaced by $\ell^2$ in flat co-ordinates, an approximation that holds for
large $\ell$, since sky curvature can be neglected for small angular
scales. We find for the flat-sky shear power spectrum
%
\begin{align}
  P^\gamma_{ij}(\ell) 
                 = & \frac 2 \pi \, \ell^4 \, \pref^2
                 \int_{0}^\infty \frac{\rm d \chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
                \int_{0}^\infty \frac{\rm d \chi^\prime}{\chi^\prime}
                \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
                %\nonumber \\
                %& \times
                \int_0^\infty \frac{{\rm d} k}{k^2} \, P_{\rm m}(k, \chi, \chi^\prime) \,
                {\rm j}_\ell(k \chi) \, {\rm j}_\ell(k \chi^\prime) .
  \label{eq:P_ell_gamma_full_der}
\end{align}
%
with flat-sky pre-factor $\ell^4$. See App.~\ref{sec:schmidt08} and
\ref{sec:giannantonio12} for discussions of alternative expressions for the
flat-sky power spectrum.

@@

To separate the $k$- and $\chi, \chi^\prime$-terms in
equation~(\ref{eq:C_ell_full}), we first approximate the matter power
cross-spectrum between two distances by the geometric mean of the two involved
distances \citep{2005PhRvD..72b3516C,2016arXiv161200770K},
%
\begin{equation}
 P_{\rm m}(k; \chi, \chi^\prime) = \sqrt{ P_{\rm m}(k; \chi) P_{\rm m}(k; \chi^\prime) } \, .
  \label{eq:P_geom_mean}
\end{equation}
%
This form is justified when considering the linear power spectrum, and
follows when inserting the linearly evolving density contrast $\hat \delta(\vec
k; \chi) = D_+(\chi) \hat \delta_0(\vec k)$ into equation~(\ref{eq:p_m}), where
$\delta_0$ is the present-day linearly extrapolated density contrast, and $D_+$
the linear growth factor with $D_+(0) = 1$. This is a good approximation
also in the non-linear case as shown in \cite{2016arXiv161200770K}.

@@

Following \cite{2008PhRvD..78l3506L} we expand to third order
%
\begin{equation}
  \lim_{\varepsilon \rightarrow 0} \int_0^\infty {\rm d} x \, {\rm e}^{-\epsilon (x - \nu)} g(x) {\rm J}_\nu(x)
  %\approx g(\nu) - \frac 1 2 g^{\prime\prime}(\nu) - \frac \nu 6 g^{\prime\prime\prime}(\nu) 
  \approx g(\nu) - \frac 1 2 \left.\frac{{\rm d}^2 g}{{\rm d} x^2}\right|_{x=\nu}
                 - \frac \nu 6 \left.\frac{{\rm d}^3 g}{{\rm d} x^3}\right|_{x=\nu},
\end{equation}
%
with $g(x) = k^{-1} f(k, \chi)$, $\chi=x/k$, and its derivatives $g^{(n)}(x) =
k^{-1-n} f^{(n)}(k, \chi)$ for a given $k$, where the derivatives of $f$
are with respect to the second argument $\chi$. In this series expansion, the
replacement of the integral with the evaluation of $g$ and its derivatives at
the maximum of the Bessel function is a good approximation of the integral if
$g$ is varying more slowly than the oscillating Bessel function.
% [Check this].
As we will show below, $f$ is a slowly varying function of the comoving
distance. In our case the projection kernel is
%
\begin{equation}
  f(k, \chi) = \sqrt{P_{\rm m}(k; \chi)} \, a^{-1}(\chi) \chi^{-3/2} q(\chi).
  \label{eq:f_LA08}
\end{equation}
%
In the tomographic case, the index $i$ is added to $q$ and $f$.
%
Replacing both distance integrals in equation~(\ref{eq:C_ell_full_Pk}) by their
Taylor-expansions around the maxima $\nu(\ell) = \ell + 1/2$ of the two Bessel
functions, which are $k \chi$ and $k \chi^\prime$, respectively, yields
%
\begin{align}
  C^\gamma_{ij}(\ell) \approx & \, \ellbar^2(\ell, 2) \, \pref^2
    \int_0^\infty \frac{{\rm d} k}{k^3} \, k^{-2}
    \left[ f_i(k, \chi) - \frac{1}{2 k^2} f_i^{\prime\prime}(k, \chi)
      - \frac{\nu(\ell)}{6 k^3} f_i^{\prime\prime\prime}(k, \chi) + \ldots \right]
    %\nonumber \\
    %& \times
    \left[ f_j(k, \chi) - \frac{1}{2 k^2} f_j^{\prime\prime}(k, \chi)
    - \frac{\nu(\ell)}{6 k^3} f_j^{\prime\prime\prime}(k, \chi) + \ldots \right] \, .
  \label{eq:C_ell_limber2_dk}
\end{align}
%
Changing the integration to $\chi = \nu/k$ and collecting terms according to their $\nu$-dependence:
%
\begin{align}
  C^\gamma_{ij}(\ell) \approx & \, C^\gamma_{{\rm L1}, ij}(\ell) + C^\gamma_{{\rm L2}, ij}(\ell)
      %\nonumber \\
    = %&
    \frac{\ellbar^2(\ell, 2)}{\nu^4(\ell)} \, \pref^2
    \int_0^\infty {\rm d} \chi \, \chi^3 \, % P_{\rm m}\left( k \right)
    \Bigg\{
    %\left\{
    (f_i f_j)(\nu(\ell)/\chi, \chi)
    %\right.
      \nonumber \\
    &
    %\left.
     - \frac 1 {\nu^2(\ell)}
    \left[ \frac{\chi^2}{2} \left( f^{}_i f^{\prime\prime}_j + f_i^{\prime\prime} f^{}_j \right)(\nu(\ell)/\chi, \chi)
     + \frac{\chi^3}{6} \left( f^{}_i f^{\prime\prime\prime}_j + f^{\prime\prime\prime}_i f^{}_j \right)(\nu(\ell)/\chi, \chi)
    \right]
    + {\cal O}(\nu^{-4})
    %\right\}
    \Bigg\}
    \, .
  \label{eq:C_ell_limber12_dr}
\end{align}
%
The first term corresponds to the well-known first-order Limber approximation
\citep{1953ApJ...117..134L,1992ApJ...388..272K}, which is widely used in
weak gravitational lensing. We retrieve the (spherical) standard expression by
inserting back the projection kernel (equation~\ref{eq:f_LA08}),
%
\begin{align}
  C^\gamma_{{\rm L1}, ij}(\ell) = & \, \frac{\ellbar^2(\ell, 2)}{\nu^4(\ell)} \, \pref^2 \int {\rm d} \chi \frac{ q_i(\chi) q_j(\chi) }{a^2(\chi)}
  P_{\rm m}\left(\frac{\nu(\ell)}{\chi}; \chi\right) . \Label{ExtL1Sph}
  \label{eq:C_ell_limber1}
\end{align}
%
In the Limber approximation, modes between structures at different epochs do
not contribute to the single line-of-sight integration.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Approximation for small $\ell$}
\label{sec:factor_D}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
